---
title: "PA1_template"
author: "Fariza"
date: "Sunday, September 20, 2015"
output: html_document
---

Module 5 : Peer Assessment 1

##Loading and preprocessing the data

####1. First we need to read .csv file. Must set the working directory to the correct folder

```{r}
library(lattice)
library(knitr)
setwd("C:/Users/Admin/Documents/RDS/M5/peerAsg1")
dfactivity<-read.csv("activity.csv")
``` 

###What is mean total number of steps taken per day?

####1. We total up all the steps for each date, change the column name, and check the first 10 lines of the total data frame totaldf

```{r}
## (total number of (steps taken per day))
totaldf <- aggregate(steps ~ date, dfactivity, sum, na.rm=TRUE)

## add descriptive variable names
names(totaldf)[2] <- "sum_steps"

## check out new data frame
head(totaldf, 5)
```

####2. Now let's see the histogram plot

The`echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r, echo=FALSE}
hist(
  totaldf$sum_steps,
  col = "red",
  main = "Frequency of Total Steps per Day",
  xlab = "Steps",
  breaks = 15
)
```

####3. Calculate and report the mean and median of the total number of steps taken per day
```{r}
mean(totaldf$sum_steps)
median(totaldf$sum_steps)
```

###What is the average daily activity pattern?

####1. Make time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
## the average number of steps taken, averaged across all days for each 5-minute interval
intervaldfna <- aggregate(steps ~ interval, dfactivity, mean, na.rm=TRUE)
head(intervaldfna)

## plot time series
plot(
  x = intervaldfna$interval,
  y = intervaldfna$steps,
  type = "l",
  main = "Time Series of the 5-Minute Interval\n and the Average Steps Taken Across All Days",
  xlab = "5-Minute Interval",
  ylab = "Average Steps Across All Days"
)

```

####2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
# Use the max() function to find the maximum number of steps
intervaldfna[intervaldfna$steps==max(intervaldfna$steps),]
```

###Inputing missing values

####1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
#Check how mana NA in data frame
allNA<-is.na(dfactivity)
sumNA<-sum(allNA)
sumNA
```

####2. Devise strategy and 3. Create new dataset with replaces NAs

```{r}
#copy to data frame, replace all NA with the mean of the steps
dfactivity.averages <- aggregate(x=list(steps=dfactivity$steps), by=list(interval=dfactivity$interval), FUN=mean, na.rm=TRUE)
df.missing <- is.na(dfactivity$steps)
table(df.missing)

#fill in the NAs
nafiller <- function(steps, interval){
  filler <- NA
  if (!is.na(steps))
    filler <- c(steps)
  else
    filler <- (dfactivity.averages[dfactivity.averages$interval==interval, "steps"])
  return(filler)
}
#replace all NA with the mean of the steps
myfill.df <- dfactivity
myfill.df$steps <- mapply(nafiller, myfill.df$steps, myfill.df$interval)

#check the new data frame and check how many NAs. Should be 0 NAs
head(myfill.df)
sum(is.na(myfill.df))
df2.missing <- is.na(myfill.df$steps)
table(df2.missing)
totalmyfill <- aggregate(steps ~ date, myfill.df, sum)
head(totalmyfill)

```

####4. Plot the histogram and calculate mean and median

```{r}
#plot histogram
hist(
  totalmyfill$steps,
  col = "blue",
  main = "Histogram of Total Steps Taken Each Day\nNA replaced",
  xlab = "Total Steps",
  breaks = 15
)
mean(totalmyfill$steps)
median(totalmyfill$steps)
```
The values of mean and median after we replaced the NAs with average value remains the same. This indicates that the strategy to fill up NAs with average value is acceptable strategy

###what are there differences in activity patterns between weekdays and weekends?

```{r}
###DATE and DAY
dayType <- function(day){
  if (day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))
    return("Weekday")
  else if (day %in% c("Saturday", "Sunday"))
    return("Weekend")
  }  
head(myfill.df)
fillDate.df <-myfill.df
head(fillDate.df)
fillDate.df$day <-weekdays(as.Date(fillDate.df$date))
fillDate.df$daytype <- sapply(fillDate.df$day, FUN=dayType)
head(fillDate.df)
```
Problem when calculating mean for each dayType